---
title: "Assigned reads summary Gehri-Larson pipeline"
author: "Peter Euclide"
date: "7/15/2020"
output: html_document
---




## View summary of DADA2 filtering paremeters


```{r sequence summary}
library(ggridges)
library(tidyverse)
library(ggpubr)
library(ggsci)
dada2_tracks <- read.csv("./tracked_dada2_filters.csv")


# find blanks/controls
toMatch_blanks <- c("BLK", "BLANK", "blank","blk", "Blank", "Blk", "FB", "fb")
toMatch_mock <- c("MOCK", "mock", "Mock")
toMatch_NTC <- c("NTC", "ntc")
toMatch_ExtNeg <- c("ExtNeg", "extneg")
field_blank <- grep(paste(toMatch_blanks, collapse = "|" ), dada2_tracks$SampleName, value = T)
mock_com <- grep(paste(toMatch_mock, collapse = "|" ), dada2_tracks$SampleName, value = T)
NTC <- grep(paste(toMatch_NTC, collapse = "|" ), dada2_tracks$SampleName, value = T)
ExtNeg <- grep(paste(toMatch_ExtNeg, collapse = "|" ), dada2_tracks$SampleName, value = T)

# create long file of just sample data
sample_tracks <- dada2_tracks %>% filter(!(SampleName %in% c(field_blank, mock_com, NTC, ExtNeg))) 

# create long file of just control data
control_tracks <- dada2_tracks %>% filter(SampleName %in% c(field_blank, mock_com, NTC, ExtNeg)) 

### make some summaries of read counts for section 1:

summary(sample_tracks)
summary(control_tracks)
summary(control_tracks %>% filter(SampleName%in% c(NTC, ExtNeg)))

## percent loss of reads
sample_tracks %>% mutate(loss = input-nonchim, ploss = (input-nonchim)/input) %>% summarise(MeanLoss = mean(loss/input), medianLoss=median(loss/input), meanploss=mean(ploss))
control_tracks %>% mutate(loss = input-nonchim, ploss = (input-nonchim)/input) %>% summarise(MeanLoss = mean(loss/input), medianLoss=median(loss/input), meanploss=mean(ploss))

```


### Figure 1

```{r}

## ggridges plot
y <- sample_tracks %>%  gather(.,"step", "reads", 2:(which(colnames(sample_tracks)=="SampleName")-1)) %>% 
  filter(step%in% c("input", "nonchim", "assignedReads"))

x <- control_tracks %>% filter(!Lake== is.na(Lake)) %>%  gather(.,"step", "reads", 2:(which(colnames(sample_tracks)=="SampleName")-1)) %>% 
  filter(step%in% c("input", "nonchim", "assignedReads")) 


lab_controls <- control_tracks %>% filter(is.na(Lake), !SampleName %in% mock_com) %>% summarise(meanInput=mean(input), meanNonchim=mean(nonchim), meanAss=mean(assignedReads))

y$step <- factor(y$step, levels = c("input" , "nonchim","assignedReads"))
x$step <- factor(x$step, levels = c("input" , "nonchim","assignedReads"))

thresh2 <- data.frame(Lake = unique(x$Lake), x =c(17, 0, 1569, 0, 0, 25,21,165,0))

ggplot(y, aes(x=reads, y = Lake, fill=step))+
  geom_density_ridges(alpha =0.4, scale =1)+
  geom_point(data=x, aes(x=reads, y = Lake, color = step), shape =17, alpha = .8)+
  lims(x=c(0,6000))+
  scale_fill_jama(labels=c("raw reads", "filtered reads", "assgined reads"))+
  scale_color_jama(labels=c("raw reads", "filtered reads", "assgined reads"))+
  theme_bw()+
  theme(text=element_text(size = 16), legend.position = "top")
#ggsave("./figures/figure_1_b.tiff", height = 10, width = 6)



  
```


## Compare known and eDNA communtities

This section Uses detection matricies which are reported as supplementary tables 2 and 3 to analyze data.  

```{r compare trad and edna}

trad <- read.csv("./TableS_2.csv")
edna_norm <- read.csv("./TableS_3.csv")

## remove hybrids from trad data

trad <- trad %>% filter(!com_name %in% grep("Hybrid", trad$com_name, value =T))


## known richness
fun=function(x){sum(x>0)}
known_rich=apply(trad[,c(5:13)], 2, fun)
#raw_rich=apply(edna_raw[,c(5:13)], 2, fun)
norm_rich=apply(edna_norm[,c(5:13)], 2, fun)

rbind(known_rich, norm_rich)

trad$data="trad"
edna_norm$data="norm"
#edna_raw$data="raw"

comb <- rbind(trad, edna_norm)

comb_long <- comb %>% pivot_longer(colnames(comb[,5:13]), "Lake") 
  

```



## Obj. 1: Intersection between eDNA species detection and known fish communities 

```{r  propotion of intersect}
#### Look at proportion of taxa that are found in both eDNA and known samples


lake <- NULL
sp_known <- NULL
sp_dna <- NULL
ge_known <- NULL
ge_dna <- NULL
fa_known <- NULL
fa_dna <- NULL

for(l in colnames(trad[5:13])){
  norm <- comb_long %>% filter(data == "norm", Lake == l, value >0)
  known <- comb_long %>% filter(data == "trad", Lake == l, value >0)
  sp_known <- c(sp_known,length(unique(known$species)))
  sp_dna <- c(sp_dna,length(intersect(unique(norm$species), unique(known$species))))
  ge_known <- c(ge_known,length(unique(known$genus)))
  ge_dna <- c(ge_dna,length(intersect(unique(norm$genus), unique(known$genus))))
  fa_known <- c(fa_known,length(unique(known$family)))
  fa_dna <- c(fa_dna,length(intersect(unique(norm$family), unique(known$family))))
  # tmp_sp <- c(length(intersect(unique(norm$species), unique(known$species))), length(unique(known$species)))
  # tmp2_sp <- c(tmp2_sp, tmp_sp)
  # tmp_gen <- length(intersect(unique(norm$genus), unique(known$genus)))/length(unique(known$genus))
  # tmp2_gen <- c(tmp2_gen, tmp_gen)
  # tmp_fam <- length(intersect(unique(norm$family), unique(known$family)))/length(unique(known$family))
  # tmp2_fam <- c(tmp2_fam, tmp_fam)
  lake <- c(lake, l)
}


## Table 2
df <- data.frame(lake, sp_known, sp_dna, ge_known, ge_dna, fa_known, fa_dna)
(df2 <- df %>% mutate(sp_per=sp_dna/sp_known, ge_per=ge_dna/ge_known, fa_per=fa_dna/fa_known))

mean(df2$sp_per)
mean(df2$ge_per)
mean(df2$fa_per)

## novel eDNA detections

lake <- NULL
sp_known <- NULL
sp_dna <- NULL
ge_known <- NULL
ge_dna <- NULL
fa_known <- NULL
fa_dna <- NULL

mybiglist <- list()
for(l in colnames(trad[5:13])){
  norm <- comb_long %>% filter(data == "norm", Lake == l, value >0)
  known <- comb_long %>% filter(data == "trad", Lake == l, value >0)
  #sp_known <- c(sp_known,length(unique(known$species)))
  sp_dna <- setdiff(unique(norm$species), unique(known$species))
 # ge_known <- c(ge_known,length(unique(known$genus)))
  ge_dna <- setdiff(unique(norm$genus), unique(known$genus))
  #fa_known <- c(fa_known,length(unique(known$family)))
  fa_dna <- setdiff(unique(norm$family), unique(known$family))
  tmp <- list(sp_dna, ge_dna, fa_dna)
  mybiglist[[l]] <- tmp
  # tmp_sp <- c(length(intersect(unique(norm$species), unique(known$species))), length(unique(known$species)))
  # tmp2_sp <- c(tmp2_sp, tmp_sp)
  # tmp_gen <- length(intersect(unique(norm$genus), unique(known$genus)))/length(unique(known$genus))
  # tmp2_gen <- c(tmp2_gen, tmp_gen)
  # tmp_fam <- length(intersect(unique(norm$family), unique(known$family)))/length(unique(known$family))
  # tmp2_fam <- c(tmp2_fam, tmp_fam)
}

## Table 2
# df1 <- data.frame(lake, sp_known, sp_dna, ge_known, ge_dna, fa_known, fa_dna)
# df1 %>% mutate(sp_per=sp_dna/sp_known, ge_per=ge_dna/ge_known, fa_per=fa_dna/fa_known)

```


```{r detection of common species}

common_sp <- comb_long %>% filter(data== "trad") %>% group_by(Lake) %>% filter(value == max(value))


lake <- NULL
sp_known <- NULL
sp_dna <- NULL
ge_known <- NULL
ge_dna <- NULL
fa_known <- NULL
fa_dna <- NULL

l="Trout"
for(l in colnames(trad[5:13])){
  norm <- comb_long %>% filter(species %in% common_sp$species, data == "norm", Lake == l, value >0)
  known <- common_sp %>% filter(species %in% common_sp$species,data == "trad", Lake == l, value >0)
  sp_known <- c(sp_known,length(unique(known$species)))
  sp_dna <- c(sp_dna,length(intersect(unique(norm$species), unique(known$species))))
  ge_known <- c(ge_known,length(unique(known$genus)))
  ge_dna <- c(ge_dna,length(intersect(unique(norm$genus), unique(known$genus))))
  fa_known <- c(fa_known,length(unique(known$family)))
  fa_dna <- c(fa_dna,length(intersect(unique(norm$family), unique(known$family))))
  lake <- c(lake, l)
}


## Table 2 commons
df <- data.frame(lake, sp_known, sp_dna, ge_known, ge_dna, fa_known, fa_dna)
(df2 <- df %>% mutate(sp_per=sp_dna/sp_known, ge_per=ge_dna/ge_known, fa_per=fa_dna/fa_known) )
mean(df2$sp_per)
mean(df2$ge_per)
mean(df2$fa_per)


```


## Obj 2. Efficiency of eDNA among systems with different known community compositions 
```{r NMDS, fig.height=10, fig.width=10}

library(vegan)


## functions:

# function to convert a detection matrix into a wide NMDS matrix ------------
lakes <- c("Crystal", "McDermott", "Mendota", "Sparkling", "Trout", "Trout_Bog" ,"UMR_13" ,"UMR_19" ,"Wingra")

det2nmds <- function(x, name){
  x <- x[c(rowSums(x[,lakes]))>0,]
  # remove any genus, or family level IDs
  x <- x %>% filter(species %in% grep(".", fixed=T, x$species, value = T)) %>% select(-c("com_name", "family", "genus", "Trout_Bog"))
  # Transpose
  x <- data.frame(t(x ))
  colnames(x) <- x[1,]
  x <- x[-1,]
  row.names(x) <- paste0(name, row.names(x))
  x
  
}



det2nmds2 <- function(x, name){
  x <- x[c(rowSums(x[,lakes]))>0,]
  # remove any genus, or family level IDs
  x <- x %>% select(-c( "Trout_Bog"))
  # Transpose
  x <- data.frame(t(x ))
  colnames(x) <- x[1,]
  x <- x[-1,]
  row.names(x) <- paste0(name, row.names(x))
  x
  
}

#  function to combine matricies --------------
combineMatrix <- function(mat1, mat2=NULL, mat3=NULL){
  comb <- bind_rows(mat1, mat2, mat3)
  rows <- rownames(comb)
  comb <- data.frame(apply(comb, 2, as.numeric))
  comb <- comb %>% mutate_if(is.numeric, ~1*(.>0))
  comb[is.na(comb)] <- 0
  rownames(comb) <- rows
  comb
}


# funciton to make biplots---------------
PlotNMDS <- function(x, title=NULL) {
  ordiplot(x,type="n", main = title)
  ordihull(x, groups=treat, draw="polygon",col=c("orange","steelblue"),label=F, alpha = .3)
  orditorp(x,display="species",col="black",air=0.01)
  orditorp(x,display="sites",col=c(rep("steelblue",8), rep("orange",8)),
         air=0.01,cex=1.25) 
}


###  Species level NMDS -----------------------------------------------


## complete known dataset ----------------
known_all <- read.csv("./TableS_3.csv")
known_all_nmds <- det2nmds(known_all, "known_all_")


## complete eDNA dataset ----------------
eDNA_all <- read.csv("./TableS_2.csv")
eDNA_all_nmds <- det2nmds(eDNA_all, "eDNA_")

###  genus level NMDS -----------------------------------------------


known_all_gen <- known_all %>% group_by(genus) %>% filter(genus!="unk")%>%summarise(Crystal=sum(Crystal),McDermott =sum(McDermott ),Mendota  =sum(Mendota ),Sparkling =sum(Sparkling ),Trout =sum(Trout ),Trout_Bog =sum(Trout_Bog ),UMR_13  =sum(UMR_13  ),UMR_19   =sum(UMR_19  ),Wingra =sum(Wingra ))
known_all_gen_nmds <- det2nmds2(known_all_gen, "known_all_")


eDNA_all_gen <- eDNA_all %>% group_by(genus) %>% filter(genus!="unk")%>%summarise(Crystal=sum(Crystal),McDermott =sum(McDermott ),Mendota  =sum(Mendota ),Sparkling =sum(Sparkling ),Trout =sum(Trout ),Trout_Bog =sum(Trout_Bog ),UMR_13  =sum(UMR_13  ),UMR_19   =sum(UMR_19  ),Wingra =sum(Wingra ))
eDNA_all_gen_nmds <- det2nmds2(eDNA_all_gen, "eDNA_all_")


###  family level NMDS -----------------------------------------------


known_all_fam <- known_all %>% group_by(family) %>% filter(family!="unk")%>%summarise(Crystal=sum(Crystal),McDermott =sum(McDermott ),Mendota  =sum(Mendota ),Sparkling =sum(Sparkling ),Trout =sum(Trout ),Trout_Bog =sum(Trout_Bog ),UMR_13  =sum(UMR_13  ),UMR_19   =sum(UMR_19  ),Wingra =sum(Wingra ))
known_all_fam_nmds <- det2nmds2(known_all_fam, "known_all_")


eDNA_all_fam <- eDNA_all %>% group_by(family) %>% filter(family!="unk")%>%summarise(Crystal=sum(Crystal),McDermott =sum(McDermott ),Mendota  =sum(Mendota ),Sparkling =sum(Sparkling ),Trout =sum(Trout ),Trout_Bog =sum(Trout_Bog ),UMR_13  =sum(UMR_13  ),UMR_19   =sum(UMR_19  ),Wingra =sum(Wingra ))
eDNA_all_fam_nmds <- det2nmds2(eDNA_all_fam, "eDNA_all_")




### create NMBDS matrix

nmds_mat_all_sp <- combineMatrix(eDNA_all_nmds, known_all_nmds)
nmds_mat_all_ge <- combineMatrix(eDNA_all_gen_nmds, known_all_gen_nmds)
nmds_mat_all_fam <- combineMatrix(eDNA_all_fam_nmds, known_all_fam_nmds)


####### Run  NMDS  ------------


comb_NMDS_sp_all=metaMDS(nmds_mat_all_sp,k=2,trymax=100)
comb_NMDS_ge_all=metaMDS(nmds_mat_all_ge,k=2,trymax=100)
comb_NMDS_fa_all=metaMDS(nmds_mat_all_fam,k=2,trymax=100)


## Calculate p value
# set treaments
treat=c(rep("eDNA", 8),
        rep("known", 8))

get_p <- function(x){
  x$grp <- treat
  ano = anosim(x[,1:(ncol(x)-2)], x$grp, distance = "bray", permutations = 9999)
  ano$signif
}

bigList <- list(nmds_mat_all_sp,nmds_mat_all_ge,nmds_mat_all_fam)

pVals <- lapply(bigList, get_p)
unlist(pVals)



# comb$grp <- c(rep("Know",7), rep("edna",7))
# ano = anosim(comb[,1:(ncol(comb)-2)], comb$grp, distance = "bray", permutations = 9999)

# goodness(comb_NMDS)
# stressplot(comb_NMDS)

## plot ordination


treat <- rep()

par(mfrow=c(3,2))

#PlotNMDS(comb_NMDS_sp_2018, "2018 catch - species")
PlotNMDS(comb_NMDS_sp_all, "2014-2018 catch - species")

#PlotNMDS(comb_NMDS_ge_2018, "2018 catch - genus")
PlotNMDS(comb_NMDS_ge_all, "2014-2018 catch - genus")

#PlotNMDS(comb_NMDS_fa_2018, "2018 catch - family")
PlotNMDS(comb_NMDS_fa_all, "2014-2018 catch - family")



```



### FIgure 2 NMDS start:

### SP NMDS

```{r NMDS ggplot - species, fig.width=8}
treat1=c(rep("eDNA", 8),
        rep("known", 8))
treat2 <- c("Oligotrophic","Mesotrophic", "Eutrophic", "Oligotrophic", "Oligotrophic", "River", "River", "Eutrophic","Oligotrophic","Mesotrophic", "Eutrophic", "Oligotrophic", "Oligotrophic", "River", "River", "Eutrophic")





data.scores <- as.data.frame(scores(comb_NMDS_sp_all))  #Using the scores function from vegan to extract the site scores and convert to a data.frame
data.scores$site <- gsub(".*_", "" ,rownames(data.scores))   # create a column of site names, from the rownames of data.scores
data.scores$site[c(6,7,14,15)] <- c("UMR-13", "UMR-19", "UMR-13", "UMR-19")
data.scores$Dataset <- treat1  #  add the grp variable created earlier
data.scores$System <- treat2  #  add the grp variable created earlier

#head(data.scores)  #look at the data



#### create ellipse
NMDS.mean=aggregate(data.scores[,1:2],list(group=treat1),mean)

# function to create ellipse
 veganCovEllipse<-function (cov, center = c(0, 0), scale = 1, npoints = 100) 
  {
    theta <- (0:npoints) * 2 * pi/npoints
    Circle <- cbind(cos(theta), sin(theta))
    t(center + scale * t(Circle %*% chol(cov)))
  }


df_ell <- data.frame()
  for(g in unique(data.scores$Dataset)){
    df_ell <- rbind(df_ell, cbind(as.data.frame(with(data.scores[data.scores$Dataset==g,],
                    veganCovEllipse(cov.wt(cbind(NMDS1,NMDS2),wt=rep(1/length(NMDS1),length(NMDS1)))$cov,center=c(mean(NMDS1),mean(NMDS2)))))
                    ,Dataset=g))
  }

## create species scores
species.scores <- as.data.frame(scores(comb_NMDS_sp_all, "species"))  #Using the scores function from vegan to extract the species scores and convert to a data.frame
species.scores$species <- rownames(species.scores)  # create a column of species, from the rownames of species.scores
#head(species.scores)  #look at the data




library(ggsci)
library(ggrepel)
sp_plot <- ggplot() + 
  #geom_text(data=species.scores,aes(x=NMDS1,y=NMDS2,label=species),alpha=0.5) +  # add the species labels
  geom_point(data=data.scores,aes(x=NMDS1,y=NMDS2,shape=System,colour=Dataset),size=3) + # add the point markers
  geom_text_repel(data=data.scores,aes(x=NMDS1,y=NMDS2,label=site),size=3,vjust=0) +  # add the site labels
  scale_color_jama()+ 
  geom_path(data=df_ell, aes(x=NMDS1, y=NMDS2,colour=Dataset), size=1, linetype=2)+
  coord_equal() +
  theme(aspect.ratio = 1)+
  theme_bw()+
  theme(plot.margin = unit(c(0,0,0,0), "cm"))

## p value

nmds_mat_all_sp$grp <- treat1
(ano = anosim(nmds_mat_all_sp[,1:(ncol(nmds_mat_all_sp)-2)], nmds_mat_all_sp$grp, distance = "bray", permutations = 9999))

```
### GE NMDS

```{r NMDS ggplot - genus,fig.width=8}

NMDS_data <- comb_NMDS_ge_all

#treat1=c(rep("eDNA", 7),
#        rep("known", 7))
#treat2 <- c("Mesotrophic", "Eutrophic", "Oligotrophic", "Oligotrophic", "River", "River", "Eutrophic","Mesotrophic", "Eutrophic", "Oligotrophic", "Oligotrophic", "River", "River", "Eutrophic")

data.scores <- as.data.frame(scores(NMDS_data))  #Using the scores function from vegan to extract the site scores and convert to a data.frame
data.scores$site <- gsub(".*_", "" ,rownames(data.scores))   # create a column of site names, from the rownames of data.scores
data.scores$site[c(6,7,14,15)] <- c("UMR-13", "UMR-19", "UMR-13", "UMR-19")
data.scores$Dataset <- treat1  #  add the grp variable created earlier
data.scores$System <- treat2  #  add the grp variable created earlier

head(data.scores)  #look at the data



#### create ellipse
NMDS.mean=aggregate(data.scores[,1:2],list(group=treat1),mean)

# function to create ellipse
 veganCovEllipse<-function (cov, center = c(0, 0), scale = 1, npoints = 100) 
  {
    theta <- (0:npoints) * 2 * pi/npoints
    Circle <- cbind(cos(theta), sin(theta))
    t(center + scale * t(Circle %*% chol(cov)))
  }


df_ell <- data.frame()
  for(g in unique(data.scores$Dataset)){
    df_ell <- rbind(df_ell, cbind(as.data.frame(with(data.scores[data.scores$Dataset==g,],
                    veganCovEllipse(cov.wt(cbind(NMDS1,NMDS2),wt=rep(1/length(NMDS1),length(NMDS1)))$cov,center=c(mean(NMDS1),mean(NMDS2)))))
                    ,Dataset=g))
  }

## create species scores
species.scores <- as.data.frame(scores(NMDS_data, "species"))  #Using the scores function from vegan to extract the species scores and convert to a data.frame
species.scores$species <- rownames(species.scores)  # create a column of species, from the rownames of species.scores
head(species.scores)  #look at the data



ge_plot <- ggplot() + 
  #geom_text(data=species.scores,aes(x=NMDS1,y=NMDS2,label=species),alpha=0.5) +  # add the species labels
  geom_point(data=data.scores,aes(x=NMDS1,y=NMDS2,shape=System,colour=Dataset),size=3) + # add the point markers
  geom_text_repel(data=data.scores,aes(x=NMDS1,y=NMDS2,label=site),size=3,vjust=0) +  # add the site labels
  scale_color_jama()+ 
  geom_path(data=df_ell, aes(x=NMDS1, y=NMDS2,colour=Dataset), size=1, linetype=2)+
  coord_fixed(ratio=1.5)+
  theme_bw()+
  theme(#plot.margin = unit(c(0,0,0,0), "cm")
    legend.position = "none")

## p value

nmds_mat_all_ge$grp <- treat1
(ano = anosim(nmds_mat_all_ge[,1:(ncol(nmds_mat_all_ge)-2)], nmds_mat_all_ge$grp, distance = "bray", permutations = 9999))


```
### FA NMDS

```{r NMDS ggplot - family,fig.width=8}

NMDS_data <- comb_NMDS_fa_all

# treat1=c(rep("eDNA", 7),
#         rep("known", 7))
# treat2 <- c("Mesotrophic", "Eutrophic", "Oligotrophic", "Oligotrophic", "River", "River", "Eutrophic","Mesotrophic", "Eutrophic", "Oligotrophic", "Oligotrophic", "River", "River", "Eutrophic")

NMDS_data$ngrp

data.scores <- as.data.frame(scores(NMDS_data))  #Using the scores function from vegan to extract the site scores and convert to a data.frame
data.scores$site <- gsub(".*_", "" ,rownames(data.scores))   # create a column of site names, from the rownames of data.scores
data.scores$site[c(6,7,14,15)] <- c("UMR-13", "UMR-19", "UMR-13", "UMR-19")
data.scores$Dataset <- treat1  #  add the grp variable created earlier
data.scores$System <- treat2  #  add the grp variable created earlier

head(data.scores)  #look at the data



#### create ellipse
NMDS.mean=aggregate(data.scores[,1:2],list(group=treat1),mean)

# function to create ellipse
 veganCovEllipse<-function (cov, center = c(0, 0), scale = 1, npoints = 100) 
  {
    theta <- (0:npoints) * 2 * pi/npoints
    Circle <- cbind(cos(theta), sin(theta))
    t(center + scale * t(Circle %*% chol(cov)))
  }


df_ell <- data.frame()
  for(g in unique(data.scores$Dataset)){
    df_ell <- rbind(df_ell, cbind(as.data.frame(with(data.scores[data.scores$Dataset==g,],
                    veganCovEllipse(cov.wt(cbind(NMDS1,NMDS2),wt=rep(1/length(NMDS1),length(NMDS1)))$cov,center=c(mean(NMDS1),mean(NMDS2)))))
                    ,Dataset=g))
  }

## create species scores
species.scores <- as.data.frame(scores(NMDS_data, "species"))  #Using the scores function from vegan to extract the species scores and convert to a data.frame
species.scores$species <- rownames(species.scores)  # create a column of species, from the rownames of species.scores
head(species.scores)  #look at the data



fa_plot <- ggplot() + 
  #geom_text(data=species.scores,aes(x=NMDS1,y=NMDS2,label=species),alpha=0.5) +  # add the species labels
  geom_point(data=data.scores,aes(x=NMDS1,y=NMDS2,shape=System,colour=Dataset),size=3) + # add the point markers
  geom_text_repel(data=data.scores,aes(x=NMDS1,y=NMDS2,label=site),size=3,vjust=0) +  # add the site labels
  scale_color_jama()+ 
  geom_path(data=df_ell, aes(x=NMDS1, y=NMDS2,colour=Dataset), size=1, linetype=2)+
  coord_equal() +
  theme_bw()+
  theme(#plot.margin = unit(c(0,0,0,0), "cm")
    legend.position = "none")


## p value

nmds_mat_all_fam$grp <- treat1
(ano = anosim(nmds_mat_all_fam[,1:(ncol(nmds_mat_all_fam)-2)], nmds_mat_all_fam$grp, distance = "bray", permutations = 9999))


```


### Figure 2: NMDS plot

```{r final figure, fig.height=10, fig.width=10}
library(ggpubr)
ggarrange(sp_plot, ge_plot, fa_plot, common.legend = T, nrow=3, ncol=1, legend = "right", align="h")


#+theme(plot.margin(t = 0, r = 0, b = 0, l = 0, unit = "pt"))


#ggsave("./figures/figure_2.tiff", height=8, width = 5)


```

## Obj 3. Relationship between life history traits and species detection 


```{r overall life historhy differences, fig.height=12 }
library(adegenet)
LH <- read.csv("TableS_4.csv")
LH[is.na(LH)] <- 0


## novel eDNA detections

lake <- NULL
sp_known <- NULL
sp_dna <- NULL
ge_known <- NULL
ge_dna <- NULL
fa_known <- NULL
fa_dna <- NULL

successfulDet <- list()
for(l in colnames(trad[5:13])){
  norm <- comb_long %>% filter(data == "norm", Lake == l, value >0)
  known <- comb_long %>% filter(data == "trad", Lake == l, value >0)
  sp_dna <- intersect(unique(norm$species), unique(known$species))
  tmp <- list(sp_dna)
  successfulDet[[l]] <- tmp

}

successfulDet_vector <- unique(unlist(successfulDet))




failedDet <- list()
for(l in colnames(trad[5:13])){
  norm <- comb_long %>% filter(data == "norm", Lake == l, value >0)
  known <- comb_long %>% filter(data == "trad", Lake == l, value >0)
  sp_dna <- setdiff(unique(known$species),unique(norm$species))
  tmp <- list(sp_dna)
  failedDet[[l]] <- tmp

}
failedDet_vector <- unique(unlist(failedDet))

## make LH matrix

success_mat <- LH %>% mutate(Scientific.Name= gsub(" ", ".", fixed=T,Scientific.Name)) %>% filter(Scientific.Name%in%successfulDet_vector) %>% mutate(grp = "successful")
failed_mat <- LH %>% mutate(Scientific.Name= gsub(" ", ".", fixed=T,Scientific.Name)) %>% filter(!Scientific.Name%in%successfulDet_vector, Scientific.Name%in%failedDet_vector) %>% mutate(grp = "failed")

dapc_data <- bind_rows(success_mat, failed_mat)
rownames(dapc_data) <- dapc_data$Scientific.Name
dapc1 <- dapc(dapc_data[,2:12],dapc_data$grp, n.pc=5, n.da=5)

scatter(dapc1, legend = T)
loadingplot(dapc1$var.contr, thresh = 0.01)

#### lake specific graphs

mk_DAPC_dat <- function(lake){
  
  sub_suc <- unlist(successfulDet[lake])
  sub_fai <- unlist(failedDet[lake])
  success_mat <- LH %>% mutate(Scientific.Name= gsub(" ", ".", fixed=T,Scientific.Name)) %>% filter(Scientific.Name%in%sub_suc) %>% mutate(grp = "successful")
  failed_mat <- LH %>% mutate(Scientific.Name= gsub(" ", ".", fixed=T,Scientific.Name)) %>% filter(Scientific.Name%in%sub_fai) %>% mutate(grp = "failed")
  dapc_data <- bind_rows(success_mat, failed_mat)
  row.names(dapc_data) <- dapc_data$Scientific.Name
  dapc(dapc_data[,2:12],as.character(dapc_data$grp), n.pc=5, n.da=2)
  
}
  
  
## ggplot version


# LD density plot
DAPC_den <- function(x,lake){
  my_df <- data.frame(x$ind.coord,grp=x$grp, spp=rownames(x$ind.coord))
  LABS <- my_df %>% filter(LD1==min(LD1) | LD1==max(LD1))
  LABS <- LABS %>% mutate(spp = paste(substr(LABS$spp, 1,1), gsub("^.*\\.","",LABS$spp), sep = ". "))
  
  ggplot()+
  geom_density(data=my_df, aes(LD1, color = as.factor(grp), fill = as.factor(grp)),alpha = 0.5)+
  geom_point(data=my_df, aes(x=LD1, y = 0, color = as.factor(grp)))+
  theme_bw()+
  scale_fill_jama(name = "",labels = c("Not Detected","Detected"))+
  scale_color_jama(name = "", labels = c("Not Detected","Detected"))+
  labs(x=" Discriminant Function 1")+
  ggrepel::geom_label_repel(data=LABS, aes(x=LD1, y = 0.0, label=spp), fill = "white", size = 3.5, nudge_y =0.15, segment.alpha=.5, label.padding=.2, direction="both")+
  ggtitle(lake)+
  theme(text = element_text(size= 14),
       legend.text = element_text(size =14))
  
}

# LD loading plot
DAPC_loading <- function(x){
  as.data.frame(x$var.contr) %>% add_rownames(var = "Habitat") %>% 
  ggplot( aes(x=Habitat, y = LD1))+
  #geom_point()+
  geom_bar(stat = "identity", width =.1)+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

}
  
  

### create lake specific plots
# overall
den0 <- DAPC_den(dapc1, "All sites")+theme(legend.position = "none")


# Crystal
dat8 <- mk_DAPC_dat("Crystal")
den8 <- DAPC_den(dat8,lake = "Crystal") + labs(x="")

lod8 <- DAPC_loading(dat8)

# McDermott
dat1 <- mk_DAPC_dat("McDermott")
den1 <- DAPC_den(dat1,lake = "McDermott") + labs(x="", y = "")
lod1 <- DAPC_loading(dat1)

# MenDota
dat2 <- mk_DAPC_dat("Mendota")
den2 <- DAPC_den(dat2, lake = "Mendota") + labs(x="")
lod2 <- DAPC_loading(dat2)

# Wingra
dat3 <- mk_DAPC_dat("Wingra")
den3 <- DAPC_den(dat3,lake = "Wingra") + labs(x="", y = "")
lod3 <- DAPC_loading(dat3)

# Sparkling
dat4 <- mk_DAPC_dat("Sparkling")
den4 <- DAPC_den(dat4,lake = "Sparkling") + labs(x="")
lod4 <- DAPC_loading(dat4)

# Trout
dat5 <- mk_DAPC_dat("Trout")
den5 <- DAPC_den(dat5,lake = "Trout") + labs(x="", y = "")
lod5 <- DAPC_loading(dat5)

# UMR-13
dat6 <- mk_DAPC_dat("UMR_13")
den6 <- DAPC_den(dat6,lake = "UMR-13")
lod6 <- DAPC_loading(dat6)

# UMR-19
dat7 <- mk_DAPC_dat("UMR_19")
den7 <- DAPC_den(dat7,lake = "UMR-19") + labs( y = "")
lod7 <- DAPC_loading(dat7)

ggarrange(den1, lod1,
          den2, lod2,
          den3, lod3,
          den4, lod4,
          den5, lod5,
          den6, lod6,
          den6, lod7,
          ncol=2, nrow=7, legend = "top", common.legend = T)



p1 <- ggarrange(den8,
          den1,
          den2,
          den3,
          den4,
          den5,
          den6,
          den7,
          ncol=2, nrow=4, legend = "none")


ggarrange(den0, p1, ncol =1,nrow=2, heights=c(1.5,4), common.legend = T)

#ggsave("./figures/figure_3.tiff", height = 12, width = 7)

temp0 <- summary(dapc1)$assign.per.pop*100
temp8 <- summary(dat8)$assign.per.pop*100
temp1 <- summary(dat1)$assign.per.pop*100
temp2 <- summary(dat2)$assign.per.pop*100
temp3 <- summary(dat3)$assign.per.pop*100
temp4 <- summary(dat4)$assign.per.pop*100
temp5 <- summary(dat5)$assign.per.pop*100
temp6 <- summary(dat6)$assign.per.pop*100
temp7 <- summary(dat7)$assign.per.pop*100

bind_rows(temp8,temp0, temp1, temp2, temp3, temp4, temp5, temp6 ,temp7)


### evaluate loading scores and plot

dat1_DF <- rownames_to_column(data.frame(dat1$var.contr), "hab" )%>% mutate(lake = "Crystal")
top_n(dat1_DF, 2,LD1)

dat2_DF <- rownames_to_column(data.frame(dat2$var.contr), "hab" )%>% mutate(lake = "McDermott")
top_n(dat2_DF, 2,LD1)

dat3_DF <- rownames_to_column(data.frame(dat3$var.contr), "hab" )%>% mutate(lake = "Wingra")
top_n(dat3_DF, 2,LD1)

dat4_DF <- rownames_to_column(data.frame(dat4$var.contr), "hab" )%>% mutate(lake = "Sparkling")
top_n(dat4_DF, 2,LD1)


dat5_DF <- rownames_to_column(data.frame(dat5$var.contr), "hab" )%>% mutate(lake = "Trout")
top_n(dat5_DF, 2,LD1)


dat6_DF <- rownames_to_column(data.frame(dat6$var.contr), "hab" )%>% mutate(lake = "UMR_13")
top_n(dat6_DF, 2,LD1)

dat7_DF <- rownames_to_column(data.frame(dat7$var.contr), "hab" ) %>% mutate(lake = "UMR_19")
top_n(dat7_DF, 2,LD1)


lods <- bind_rows(dat1_DF,
          dat2_DF,
          dat3_DF,
          dat4_DF,
          dat5_DF,
          dat6_DF,
          dat7_DF)


lods$hab <- as.factor(lods$hab)
levels(lods$hab) <- c("Bedrock", "Boulder", "Clay/Silt", "Cobble", "Gravel", "Woody Debris", "Muck", "Organic Debris", "Pelagic", "Prefer Lotic", "Sand", "Vegetation")

ggplot(lods, aes(x = hab, y = LD1, fill = lake))+
  geom_bar(stat= "identity", position="dodge")+
  scale_fill_jama()+
  theme_bw()+
  labs(fill = "", x = "Habitat Preference", y="Variable Contribution")+
  theme(text = element_text(size = 12),
        legend.position = c(0.1,0.75))

#ggsave("./figures/figure_4.tiff", width = 10, height = 5 )
```


